---
title: "week_7"
author: "Veronica Thompson"
date: "September 3, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```


### Chapter 13 - Relational data  

* Read 13.1 - 13.7  
* 13.2.1 Problems 1 - 4  
* 13.3.1 Problems 1, 2.1, 2.2 (install the correct packages first!)  
* 13.4.6 Problems 1 and 2  
* 13.5.1 Problems 2 and 4  


Tools for using data from multiple related tables  
*Mutating joins*  
*Filtering joins*  
*Set operations*  


The airlines data sets:  
```
airlines #airline carrier names by abbreviation 
airports #airport info, duh
planes #plane by tailnum
weather #at NYC airports per hour
```

#### 13.2.1 Exercises  

*1. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?*  

from flights, you need the origin and destination airport  
from airports, you will need lat and lon for each airport matched by FAA  
You would use a mutate join!  

*2. I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?*  

```
weather$origin == airports$faa 
```  

*3. weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?*  

weather would also match the destination in flights  

*4. We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?*  

table Special would include the following variables:  
* Year  
* Month  
* Day (connected to the same variables in flights and weather)  
* number of flights incoming to NYC 
* number of flights outgoing from NYC 

####13.3.1 Exercises  

*1. Add a surrogate key to flights.*  
```{r}
flights2 = flights

flights2$surr_key = 1:nrow(flights2)
head(flights2)
```

*2. Identify the keys in the following datasets:*  
*Lahman::Batting*  
```{r}
library(Lahman)
head(Batting)

Batting %>% count(playerID, yearID, stint) %>% filter(n>1) 
#this is dumb
```  

*babynames::babynames*  
```{r}
library(babynames)
head(babynames)
babynames %>% count(year, name, sex) %>% filter(n>1)
```  

#### 13.4.6 Exercises  

*1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:*  

```{r}
flights2 <- flights %>% 
  group_by(dest) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE))
head(flights2)
```

```{r}
airports %>%
  right_join(flights2, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat, color = delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```  

*2. Add the location of the origin and destination (i.e. the lat and lon) to flights*  
```{r}
flights_loc1 <- flights %>%
  left_join(airports, by = c("origin" = "faa")) %>%
  left_join(airports, by = c("dest" = "faa")) %>%
  view()

#then select and clean this up
```  

#### 13.5.1 Exercises  

*2. Filter flights to only show flights with planes that have flown at least 100 flights.*  
```{r}
flights_count = flights %>%
  count(tailnum) %>%
  filter(n>100) 

flights_100plus = flights %>%
  semi_join(flights_count, by = "tailnum")

dim(flights)
dim(flights_100plus)
#theres less!
```  

*4. Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?*  


