---
title: "week_3"
author: "Veronica Thompson"
date: "July 18, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Chapter 4

### 4.4 Practice problems

#### 4.4.1Keyboard Shortcut Help
Why does this code not work?

```
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

Look carefully! (This may seem like an exercise in pointlessness, but training your brain to notice even the tiniest difference will pay off when programming.)

** The same variable name wasn't used (not sure what was used in the place of "i") **

#### 4.4.2
Tweak each of the following R commands so that they run correctly:

```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
# not sure what is wrong with this one...
```

```{r}
#fliter(mpg, cyl = 8), command name spelled wrong, "==" in place of "=" for logical "is equal too"
filter(mpg, cyl == 8)
```

```{r}
#filter(diamond, carat > 3), the data set is called "diamonds", not "diamond" 
filter(diamonds, carat >3)
```

#### 4.4.3
Press Alt + Shift + K. What happens? How can you get to the same place using the menus?
Alt + Shift + K brings up the keyboard shortcut guide for R! This can also be found under **Tools > Keyboard Shortcut Help**

## Cham