---
title: "week_3"
author: "Veronica Thompson"
date: "July 18, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Chapter 4  

### 4.4 Practice problems  

#### 4.4.1Keyboard Shortcut Help  
Why does this code not work?  

```
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

Look carefully! (This may seem like an exercise in pointlessness, but training your brain to notice even the tiniest difference will pay off when programming.)  

** The same variable name wasn't used (not sure what was used in the place of "i") **  

#### 4.4.2
Tweak each of the following R commands so that they run correctly:  
```{r}
library(tidyverse)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
# not sure what is wrong with this one...
```

```{r}
#fliter(mpg, cyl = 8), command name spelled wrong, "==" in place of "=" for logical "is equal too"
filter(mpg, cyl == 8)
```

```{r}
#filter(diamond, carat > 3), the data set is called "diamonds", not "diamond" 
filter(diamonds, carat >3)
```

#### 4.4.3  
Press Alt + Shift + K. What happens? How can you get to the same place using the menus?  
Alt + Shift + K brings up the keyboard shortcut guide for R! This can also be found under **Tools > Keyboard Shortcut Help**

## Chapter 5  

Assignment:  
* Read 5.1 - 5.7  
* Complete the following problems  
  + 5.2.4 Problems 1.1, 1.4, 1.5, 1.7, 2 and 3  
  + 5.3.1 Problems 1-4  
  + 5.4.1 Problems 1-4  
  + 5.5.2 Problems 1-5  
  + 5.6.7 Problem 5  
  
### 5.2

logical operators  
```{r}
#two different ways to filter for multiple values in the same variable
nov_dec1 <- filter(flights, month == 11 | month ==12)
nov_dec2 <- filter(flights, month %in% c(11,12))
```

### 5.5.4 Exercise  

for reference:  
```{r}
colnames(flights)
```

#### 5.2.4.1.1 
Had an arrival delay of two or more hours  
```{r}
ex1.1 <- filter(flights, arr_delay >= 120) #arr_delay is in minutes
summary(ex1.1$arr_delay) # arr_delay minimum is 120
```

#### 5.2.4.1.4  
Departed in summer (July, August, and September)  
```{r}
ex1.4.1 = filter(flights, month == 7 | month == 8 | month == 9)
summary(ex1.4.1$month) # month range is 7-9
```

```{r}
#alternate method
ex1.4.2 = filter(flights, month %in% c(7:9))
summary(ex1.4.2$month)
```

#### 5.2.4.1.5  
Arrived more than two hours late, but didn’t leave late  
```{r}
ex1.5 = filter(flights, arr_delay > 120 & dep_delay <= 0)
summary(ex1.5) #looks good
```

#### 5.2.4.1.7  
Departed between midnight and 6am (inclusive)  
```{r}
summary(flights$dep_time) #HMM or HHMM format, so I am looking for 2400 (midnight) and before 600 (6 am)
```

```{r}
ex1.7 <- filter(flights, dep_time == 2400 | dep_time <= 600 )
ggplot(ex1.7, aes(x= dep_time)) + geom_histogram(binwidth = 30) # looks good!
```

#### 5.2.4.2  
Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?  

From between() help file "This is a shortcut for x >= left & x <= right". For departed in summer problem  
```{r}
ex_summer = filter(flights, between(month, 7, 9))
summary(ex_summer$month)
```


#### 5.2.4.3  
How many flights have a missing dep_time? What other variables are missing? What might these rows represent?  
```{r}
no_depart = filter(flights, is.na(dep_time))
head(no_depart)
summary(no_depart)
```
It looks like there are NAs for dep_time, dep_delay, arr_time, arr_delay, air_time. These seem like cancelled flights.  

### 5.3 arrange()  

#### 5.3.1.1   
How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
first: check the help file for arrange. it was not helpful!  
```{r}
na_first = arrange(flights, desc(is.na(dep_time)))
dim(flights) == dim(na_first) #same dimensions
head(na_first) #header looks good
```

#### 5.3.1.2  
Sort flights to find the most delayed flights. Find the flights that left earliest.    
```{r}
#most delayed departure
arrange(flights, desc(dep_delay)) %>%
  head()
```
```{r}
#Earliest departure
arrange(flights, dep_delay) %>%
  head() #leaving before they are scheduled!
```

#### 5.3.1.3  
Sort flights to find the fastest flights.  
```{r}
arrange(flights, air_time) %>%
  head()
```

#### 5.3.1.4  
Which flights travelled the longest? Which travelled the shortest?  
```{r}
#shortest distance
arrange(flights, distance) %>%
  head()
```

```{r}
#longest distance
arrange(flights, desc(distance)) %>%
  head() #vacationers got a looooong flight
```

### 5.4 select()  

#### 5.4.1.1  
Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.  
```
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, starts_with("dep"), starts_with("arr"))
select(flights, ends_with("time"), ends_with("delay"), - air_time, - starts_with("sched"))
select(flights, 4, 6, 7, 9)
```

#### 5.4.1.2
What happens if you include the name of a variable multiple times in a select() call?  
Guess: it will select it twice
```{r}
head(select(flights, carrier, flight, carrier))
# only grabs it the first time it is called.
```

#### 5.4.1.3  
What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars)) #selects what is in the vector. Seems to be the same as just using "vars".
#select(flights, vars)
```

#### 5.4.1.4  
Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?  
```{r}
select(flights, contains("TIME"))
```
Selected everything that has time in the name, regardless of case. ignore.case is TRUE by default, can be set to FALSE to prevent non-specific matches by case.  
```{r}
select(flights, contains("TIME", ignore.case = FALSE))
```

### 5.5 mutate()  

#### 5.5.2.1
Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
head(transmute(flights,
                 dep_time,
                 dep_minutes = dep_time %/% 100 *60 + dep_time %% 100,
                 sched_dep_minutes = sched_dep_time %/% 100 *60 + sched_dep_time %% 100))
```

#### 5.5.2.2
Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?  
```{r}
head(transmute(flights, 
               air_time,
               arr_dep = arr_time - dep_time))
```
arr_time and dep_time need to be converted to minutes before calculating, as above.  

#### 5.5.2.3  
Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?  
```{r}
head(transmute(flights,
               dep_time,
               calc_dep_time = sched_dep_time + dep_delay,
               sched_dep_time,
               dep_delay))
```
This should work for delays under an hour. dep_time and sched_dep_time must be converted as above before calculating calc_dep_time for flights delayed over an hour.  

#### 5.5.2.4
Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().  
```{r}
delay_rank = transmute(flights, 
               flight,
               delay_ranking = min_rank(desc(dep_delay)),
               dep_delay) %>%
  arrange(delay_ranking)
head(delay_rank, 10)
```
Adjust what happens in a tie using (ties.method =  )  

#### 5.5.2.5  
What does 1:3 + 1:10 return? Why?  
```
1:3 + 1:10
#error: longer object length is not a multiple of shorter object length [1]  2  4  6  5  7  9  8 10 12 11
```  
R is unsure how to calculate when two vectors do not match in length or multiples of a length.  

### 5.6 summarise()   

#### 5.6.7.5
Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n())).  
```{r}
flights %>%
  group_by(carrier) %>%
  summarize(ave_delay = mean(dep_delay, na.rm = TRUE), n= n()) %>%
  filter(n >= 10000) %>%
  arrange(desc(ave_delay))
```

