---
title: "week_5"
author: "Veronica Thompson"
date: "August 8, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Week 5 - R4DS Data Wrangling I (VT)  
#### August 8th/9th  
#### R4DS Chapter 9 - Introduction  
#### R4DS Chapter 10 - Tibbles  
* Read 10.1 - 10.4  
* 10.5 Problems 1-2, 4-5  

## Chapter 10 - Tibbles  

### 10.5 Practice problems  

#### 10.5.1 How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).  
```{r}
print(mtcars)
```  

```{r}
mtcars_tib <- as_tibble(mtcars)
print(mtcars_tib)
```  

There are no row names in the tibble. This is only a problem if you name your rows (so don't do that??)  

#### 10.5.2 Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?  

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x # matches to column named "xyz"
df[, "xyz"] # prints as a vector, lists factor levels
df[, c("abc", "xyz")]
```

```{r}
df_tib <- as_tibble(df)
df_tib$x # does not complete partial match, as in df
df_tib[, "xyz"] #prints as a data frame
df_tib[, c("abc", "xyz")] #idk if there is a difference
```

#### 10.5.4 Practice referring to non-syntactic names in the following data frame by:  
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
annoying
```  

Extracting the variable called 1.  
```{r}
annoying$`1`
# or annoying$'1'
# or annoying$"1"
```  

Plotting a scatterplot of 1 vs 2.  
```{r}
ggplot(data = annoying, aes(x = `1`, y = `2`)) +
  geom_point()
```  

Creating a new column called 3 which is 2 divided by 1.  
```{r}
annoying %>%
  mutate(`3` = `2`/`1`)
```  

Renaming the columns to one, two and three.  
```{r}
#annoying %>%
#  rename(`1` = hi)
```  
This is surprisingly hard...  

#### 10.5.5 What does tibble::enframe() do? When might you use it?  
enframe makes a tibble with a name column, might be useful if you have a df with column names?  
```{r}
#from help file
enframe(1:3)
enframe(list(one = 1, two = 2:3, three = 4:6))
```

## Chapter 11 - Data import  
* Read 11.1 - 11.6  
* 11.2.2 Problems 1, 4 - 5  
* 11.3.5 Problems 1-3, 7  

### 11.2.2 Exercises  

#### 11.2.2.1 What function would you use to read a file where fields were separated with
“|”?  
```{r}
read_delim("a|b \n 1.0|2.0", delim = "|")
```  

#### 11.2.2.4 Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you’ll need to use read_delim() instead. What arguments do you need to specify to read the following text into a data frame?  
```{r}
read_delim("x,y \n 1,'a,b'", delim = ",", quote = "'") 
```  

#### 11.2.2.5 Identify what is wrong with each of the following inline CSV files. What happens when you run the code?  
```{r}
read_csv("a,b \n 1,2,3 \n 4,5,6") 
#there is no third column name so the thrid value in each row string is dropped
```  
```{r}
read_csv("a,b,c \n 1,2 \n 1,2,3,4") 
# row length does not match the coumn length. If too long the extra variables are dropped. If too short they are filled with NAs
``` 
```{r}
read_csv("a,b\n\"1")
#not enough varables to fill the columns, so b is left as NA
```  
```{r}
read_csv("a,b\n1,2\na,b")
#not sure what is wrong here...
```  
```{r}
read_csv("a;b\n1;3")
#wrong delim! use read_csv2()?
```  

### 11.3.5 Exercises  

#### 11.3.5.1 What are the most important arguments to 'locale()'?  
it depends on what you are trying to parse! see the help file.   

#### 11.3.5.2

#### 11.3.5.3  

#### 11.3.5.7 